The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Correction to final part of the paper
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1749 4525493e-7705-40b1-a816-d608a930855b

\section{Data Aggregation System\label{DAS}}
The goal of the Data Aggregation System was to combine existing data-services
under one umrella, utilizing their access rules, APIs, security policies, etc., and
allowing users to perform queries against them. In contrast, the approaches
discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
enforced a publication step of data into a dedicated database 
or central repositories \cite{iRODS}. Quite often it is required
that data should be converted into uniform data formats in order enable
aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
by existing data-services and infrastructure, whose design, implementation and
security policies are left up to their developers. Therefore DAS does not require
a data preservation and transaction capabilities. Instead DAS can be treated
as proxy or caching layer where information retrieval are triggered dynamically
by user queries. Such infrastructure allows data sharing across different queries and
data aggregation using appropriate mapping. Below we discuss in details the
DAS architecture.

\subsection{DAS architecture}
The design of the DAS system was based on previous studies of CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. Teh DBS Data Discovery service
\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provides
ability to place queries, similar to SQL, without specific knowledge of 
underlying DB schema.
%We carefully analyzed user
%requests and queries placed into the DBS, their patterns, frequencies and latency. 
%As a result we created a data discovery service \cite{DD}. Originally it used MVC
%architecture and later, being re-factored as pure presentation layer on top of 
%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
%we took away relations and details on underlying schema from end-users. 
%This was achieved by using Dijkstras shortest path algorithm 
%to establish necessary table joins based on provided information in user queries and
%knowledge of foreign key relationships from auto-loaded database schema.
%The DBS-QL keys exposed to end-users were choosen based on a jargon used by
%physicists. They're internally mapped into database entitites and attributes, 
%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
A quick adoption and wide usage of DBS-QL in CMS gave us confidence in choosen approach 
and provide a basis for building DAS system.

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support of keyword based search queries with ability to use condition operators;
\item pluggable interface to existing and up-coming data-service APIs;
\item support heterogeneous software environment and distributed nature of the data-services;
\item preserve security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementations and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query where placed into analytics DB and mapped into set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache users were able 
to see the results. This on demand system was supplemented by DAS robots 
(unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

@@

\subsection{DAS caching system}
The DAS caching system was used to dynamically fetch
and aggregate data upon user requests. We used the following 
algorithm:
\begin{enumerate}[1.]
\item Decompose query into set of conditions ($conds$) 
and selection keys ($skeys$)
\item Check cache for query hash, if found: look-up results from the cache
\item Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r(url, api, args) for r in mapping_db(skeys, conds)]
\end{verbatim}
Here \verb+mappind_db+ defines translation between input query and data-service
APIs which can delivery such data.
\item Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
Here \verb+superset_in_analytics+ identifies if provided \verb+api, args+
are subset of already registered in Analytics DB api call(s). The 
\verb+get_data+ and \verb+parser+ are define method to fetch and parse
data, respectively.
\item Perform aggregation by iterating over records from the previos step
\begin{verbatim}
if cache_has(row.match_key):
   existing_record = get_from_cache(row.primary_key)
   merge(row, existing_record)
   insert_into_cache(row)
   delete_from_cache(existing_record)
else:
   insert_into_cache(row)
\end{verbatim}
We did use a bulk insert operations to increase throughput. The \verb+primary_key+
is a key which identify the record defined in Mapping DB, e.g. for 
\verb+file+ record it is a \verb+file.name+.
\item Get results from the cache
\end{enumerate}
The data was stored in a form of DAS records into the cache, while
queries were recorded into Analytics DB for further analysis. We
also had an option to run independent set of daemons, DAS robots, who
monitor Analytics DB for most popular requests and pre-fetch the most
common data into DAS cache. 
%The interesting part was being able
%to identify the superset of api/args out of found one in an Analytics DB.

@@

\subsection{DAS queries}
Apart from previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys known to DAS to describe common
entities of our data. For example, {\it dataset, file, block, run, event}. Those
names were used by our users in their daily work operations. So, by typing
{\it file} we didn't perform a string matching to ``file'' string, but rather
look-up all file objects in a DAS. Moreover each DAS key has a default attribute.
For intance, for {\it dataset, block, file} the default attribute 
was a {\it name}, while for {\it run} and {\it event} the default attribute was 
a {\it number}. This allows to specify key-value pair as simple as possible, e.g.
{\it run=123456}, which means run number 123456. At the same time a query 
key, e.g. {\it site} meant to look-up all site objects. Such ability
to use key-value pairs resolve ambiguity in object values. For instance,
specifying a run number as {\it run=123456} instruct system to resolve
data object with run equal to 123456 rather then other objects which may have
123456 values, e.g. event number or file size. At the same time, to complement
the stregth of the DAS we do plan to index data objects and add a keyword-based 
search later.

The free text-based queries were transformed into MongoDB query syntax, 
who was used internally by DAS. It represents a set of selection 
keys and conditions as a JSON document. For example, a query
{\it dataset site=T1\_CH\_CERN}
where transformed into 
{\it \{"fields" : ["dataset"], "spec" : \{"site" : "T1\_CH\_CERN"\}\}}.
Such syntax was perfect fit for us, since it allows to store all queries
as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
MongoDB query syntax is very reach. It supports advanced queries with 
boolean expressions, conditional operators, regular expressions, etc. providing
a real power for our end-users.

\subsection{Data-services and aggregation}
As we outlined above the DAS Mapping DB becomes an authorative
source of information about all data-services participating in DAS.
Based on information stored in Mapping DB we were able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) and data location system (PhEDEx)
provided information about CMS data files. In former case, DBS 
stored information about files and their relationship to other 
physics objects, such as run and luminosity, while PhEDEx kept
information about the file location, e.g. details about sites 
holding copy of the files. Upon a user request to find a file, 
we were able to identify appropriate set of DBS and PhEDEx APIs, 
retrieve data from them, re-format their output to DAS notations 
and store results into DAS cache performing aggregation of matched 
objects. For instance, since both data-services returns in their
output a file name we used as a common key during aggregation.
This merged results represented a set of DAS records. Each DAS
record contains a standard DAS header, which identify look-up time,
request url, input parameters and API method as well as response 
version, expiration time and checksum of the output. Usage of this 
information was valuable contribution in debugging process of
data-services themselves, e.g. identification of not-matched records, 
latency studies, etc. Here is an example of DAS record representing
aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
and {\it CMStoSiteName}, with API results merged under the {\it site} key.
\begin{verbatim}
{"site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
 "das": [{"ctime": 0.9, "expire": 1256362175, "api": ["CMStoSAMName"],
          "url": "https://a.b.com/sitedb/json/index/CMStoSAMName", 
          "timestamp": 1256318975, "system": ["sitedb"]},
         {"ctime": 1.2, "api": ["CMStoSiteName"], "expire": 1256362174},
          "url": "https://a.b.com/sitedb/json/index/CMStoSiteName", 
          "timestamp": 1256318974, "system": ["sitedb"]]}
\end{verbatim}

@@

%Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
%\hline
%tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
%indexies (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
%\hline
%\hline
%\end{tabular}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{table*}
At the moment we don't have exact numbers for amount of data which will stored 
in those data-service and therefor accessible in DAS. Although we project to
accumulate of the order of 500 GB of meta-data per year during normal LHC and CMS
operations.
%\footnote{This number is based on tests 
%performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have signicant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

To test DAS performance we measured elapsed time and CPU/RAM utilization
required to collect and aggregate block object information from DBS and 
PhEDEx CMS data-services. Preliminary results of those measurements can
be found in Table \ref{DAS_benchmark}.

\begin{table*}[hbt]
\centering
\begin{tabular}{llllll}\hline
\hline
System & Format & Size & Records & Elapsed time & Elapsed time \\
& & & & no cached data & w/ cached data \\
\hline
DAS (DBS) & XML & 187MB & 368716 & 53 sec & 0.92 sec \\
DAS (PhEDEx) & XML & 164MB & 175884/170131 & 160 sec & 0.94 sec \\
DAS (aggregate) & JSON & 357MB & 374468 & 256 sec & 44.8 sec \\
\hline
\hline
\end{tabular}
\caption{Time required to fetch, parse and aggregate block information
from DBS and PhEDEx systems via their APIs. In the case of PhEDEx
system we show total number of fetched records 175884 together with
number of records merged in DAS 170131. The total number of DAS records 
were calculated as number of aggregated records plus left over, non-matched records,
from both systems.}
\label{DAS_benchmark}
\end{table*}

Please note that initial DAS implementation was done in Python 
and all tests were performed on x64 Linux platform with 16GB of RAM where
1 CPU was used by DAS server and 1 CPU by MongoDB database server. The CPU utilization
of DAS server was 67\% in average and RAM usage was below 35MB for
all tests. All tests were performed multiple times to avoid fluctuations. 
The elapsed time measured during those tests consists of the 
following componenets:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here the {\it retrieval time} is a time required to access data from remote data-service,
{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
is a time required to convert notations used by data-service to DAS ones for every object
we parse, {\it cache insertion} and {\it indexing time} represents time spend to add objects into
the DAS cache, {\it aggregation time} is required to merge objects into DAS records based
on their common key (block name in this case) and {\it output creation time}
is a time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, were only applied to
the one of the sub-system (either DBS or PhEDEx) since the first one was injected into
fresh database, while the second required the merging step.

@@

The individual tests of DAS components (DBS and PhEDEx) shown roughly 
the same performance. Although the time of PhEDEx component was bigger 
due to longer retrieval and parsing time.
The time of PhEDEx component shown in Table \ref{DAS_benchmark} divided 
among: retrieval + parsing time $\sim$110 sec and merging step $\sim$50 sec.
As you can see from Table \ref{DAS_benchmark},
the time spent to access individual components (last column) was quite
reasonable, roughly 1 second for both systems. While final time to
get DAS records on disk was about 45 seconds. This was mostly due to I/O and
conversion between binary data format used by MongoDB, BSON, to DAS
data format, JSON.

Based on this results we measured that we can achieve 7000 docs/sec rate
for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for reading operation
in DAS, which is suitable for our needs. Since all operations were performed 
on a single CPU core a further improvements are expected. Also, a necesarry
tuning can be done by using native C++ MongoDB driver instead of python one.

\section{Summary}
We presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
aggregation layer on top of the existing relational and non-relation data-services
mostly in real time fashion. All of the data were retrieved on demand basis,
while data pre-fetching of most common queries is in our plans. We developed
prototype in python for DAS system which is under commissioning phase right now 
and performed performance studies shown in this paper. 
The CMS experiment is started to collect its data in December of 2009 and 
we expect to accumulate a few PB of real data each year to tape. 
In addition, the Monte-Carlo samples will be produced at this scale.
Based on test studies performed in CMS, we expect that total size of
meta-data produced by experiment each year will be of the order of
$\sim500$GB. The DAS should sustain such load and will provide generic 
``data-discovery'' service for CMS experiment in years to come.

@@

%\bibitem{Python} Python programming language, http://www.python.org

>>>>

\section{Data Aggregation System\label{DAS}}
The goal of the Data Aggregation System was to combine and
aggregate meta-data from existing data-services
under one umrella, utilizing their access rules, APIs, security policies, etc., and
allowing users to perform queries against them. In contrast, the approaches
discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
enforced a publication step of data into a dedicated database 
or central repositories \cite{iRODS}. Therefore, quite often it is required
that data should be converted into uniform data formats in order enable
aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
by existing data-services and infrastructure, whose implementation,
data representation and security policies are left intact. Consequently,
DAS does not require a data preservation and transaction capabilities. 
Instead DAS can be treated
as proxy or caching layer where information retrieval are triggered dynamically
by user queries. Such infrastructure allows data sharing across different queries and
data aggregation using appropriate mapping. Below we discuss in details the
DAS architecture.

\subsection{DAS architecture}
The design of the DAS system was based on previous studies of the CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
a syntax similar to SQL, without specific knowledge of underlying DB schema.
%We carefully analyzed user
%requests and queries placed into the DBS, their patterns, frequencies and latency. 
%As a result we created a data discovery service \cite{DD}. Originally it used MVC
%architecture and later, being re-factored as pure presentation layer on top of 
%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
%we took away relations and details on underlying schema from end-users. 
%This was achieved by using Dijkstras shortest path algorithm 
%to establish necessary table joins based on provided information in user queries and
%knowledge of foreign key relationships from auto-loaded database schema.
%The DBS-QL keys exposed to end-users were choosen based on a jargon used by
%physicists. They're internally mapped into database entitites and attributes, 
%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
A quick adoption and wide usage of DBS-QL in CMS gave us confidence in choosen approach 
and provide a basis for building DAS system.

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support a keyword based search queries with ability to use condition operators;
%\item pluggable interface to existing and up-coming data-service APIs;
\item use existing heterogeneous software environment and distributed nature of the data-services;
\item preserve access rules and security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementation and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query were placed into analytics DB and mapped into a set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to the users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache, users were able 
to see the results. This on demand system was supplemented by DAS robots 
(unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

@@

\subsection{DAS caching system}
The DAS caching system was used to dynamically fetch
and aggregate data upon user requests. We used the following 
algorithm:
\begin{enumerate}[1.]
\item Decompose query into set of conditions (\verb+conds+) 
and selection keys (\verb+skeys+)
\item Check cache for query hash, if found, look-up results from the cache
\item Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r(url, api, args) for r in mapping_db(skeys, conds)]
\end{verbatim}
Here \verb+mappind_db+ defines translation between input query and data-service
APIs which can delivery such data.
\item Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
Here \verb+superset_in_analytics+ verifies if provided \verb+api, args+
are subset of already registered in Analytics DB api call(s). The 
\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
data, respectively.
\item Perform aggregation by iterating over records from the previos step
\begin{verbatim}
if cache_has(row.primary_key):
   existing_record = get_from_cache(row.primary_key)
   merge(row, existing_record)
   insert_into_cache(row)
   delete_from_cache(existing_record)
else:
   insert_into_cache(row)
\end{verbatim}
The \verb+primary_key+
is a key which identify the record defined in Mapping DB, e.g. \verb+file.name+
for \verb+file+ record.
We did use a bulk insert operations at this step to increase throughput. 
\item Get results from the cache
\end{enumerate}
%The data was stored in a form of DAS records into the cache, while
%queries were recorded into Analytics DB for further analysis.
We also had an option to run independent set of daemons, DAS robots, who
monitor Analytics DB for most popular requests and pre-fetch the most
common data into DAS cache. 
%The interesting part was being able
%to identify the superset of api/args out of found one in an Analytics DB.

@@

\subsection{DAS queries}
Apart from the previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys registered in DAS to describe common
entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
was based on analysis of daily work operations of our users, see \cite{DBS07}.
Each key had a set of attributes with primary one assigned to the key itself.
For intance, the \verb+block+ key had the following attributes
\verb+name+, \verb+size+, \verb+replica+ etc., 
with primary attribute being \verb+name+.
The choice of attributes were based on a set of meta-data information
associated with a given DAS key. Therefore end-users where able to specify
a context in their queries by using appropriate attribute(s). For example,
for query \verb+block+ we didn't perform a string matching to the
``block'' string, but rather look-up a block objects in a DAS, while query
\verb+block.name=abc*+ instructs system to look-up appropriate blocks
with matched names. At the same time, to complement the stregth of the 
DAS, we do plan to index dynamically retrieved meta-data and add a keyword-based 
search later.

The free text-based queries were transformed into MongoDB query syntax, 
who was used internally in DAS. It represents a set of selection 
keys and conditions as a JSON document. For example, a query
{\it dataset site=T1\_CH\_CERN}
was transformed into 
{\it \{"fields" : ["dataset"], "spec" : \{"site" : "T1\_CH\_CERN"\}\}}.
Such syntax was perfect fit for us, since it allows to store all queries
as JSON documents into Analytics DB for further analysis. It is worthwhile to note that
MongoDB query syntax is very reach. It supports advanced queries with 
boolean expressions, conditional operators, regular expressions, etc. providing
a real power for our end-users.

\subsection{Data-services and aggregation}
As we outlined above the DAS Mapping DB becomes an authorative
source of information about all data-services participating in DAS.
Based on information stored in Mapping DB we were able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) and data location system (PhEDEx)
provided information about CMS block meta-data. In former case, DBS 
stored information about blocks and their relationships to other 
physics objects, such as dataset and files, while PhEDEx kept
information about its location and replicas. Upon a user request to 
find a block meta-data, we were able to identify appropriate set of DBS and PhEDEx APIs, 
retrieve data from them, re-format their output to DAS notations 
and store results into DAS cache performing aggregation of matched 
objects using \verb+block.name+ key. The merged results were stored 
as a set of data records. Here is an example of such aggregated record:
\begin{verbatim}
{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
 "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
            "nfiles": "18", "se": "srm.site.com", 
            "size": "112694171838"}, {....}]},
           {"name": "a/b/c#123", "nevents": "570200", 
            "created_by": "USER_DN", "path": "/a/b/c", ...}],
 "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
\end{verbatim}
Here \verb+das_id+ refers to api call records stored in Analytics DB,
\verb+block+ represents an aggregation block meta-data information from DBS
and PhEDEx systems and \verb+das+ contains their expiration time stamps.
Usage of this information was valuable contribution in debugging process of
data-services themselves, e.g. identification of not-matched records, 
latency studies, etc.
% Here is an example of DAS record representing
%aggregation information between two SiteDB APIs, {\it CMStoSAMName} 
%and {\it CMStoSiteName}, with API results merged under the {\it site} key.
%\begin{verbatim}
%{"site":[{"samname": "CERN-PROD", "name": "T1_CH_CERN"}, 
%         {"sitename": "CERN", "name": "T1_CH_CERN"}], 
% "das": [{"ctime": 0.9, "expire": 1256362175, "api": ["CMStoSAMName"],
%          "url": "https://a.b.com/sitedb/json/index/CMStoSAMName", 
%          "timestamp": 1256318975, "system": ["sitedb"]},
%         {"ctime": 1.2, "api": ["CMStoSiteName"], "expire": 1256362174},
%          "url": "https://a.b.com/sitedb/json/index/CMStoSiteName", 
%          "timestamp": 1256318974, "system": ["sitedb"]]}
%\end{verbatim}

@@

%Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
%\hline
%tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
%indexies (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
%\hline
%\hline
%\end{tabular}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{table*}
At the moment, the amount of meta-data stored in all services is equal to
$\sim$75 GB and $\sim$200 GB in tables and index sizes, respectively. While 
we estimate to accumulate $\sim$500 GB of the meta-data each year during 
normal CMS operation at LHC.
%\footnote{This number is based on tests 
%performed in CMS during Cosmic Run at Full Tesla (CRAFT) data taking \cite{CRAFT09}.}
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have signicant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

To test DAS performance we measured elapsed time and CPU/RAM utilization
required to collect and aggregate information from different
data-services. Table \ref{DAS_benchmark} summarizes the biggest 
contributors so far, the block meta-data information provided by DBS and 
PhEDEx CMS data-services.

\begin{table*}[hbt]
\centering
\begin{tabular}{llllll}\hline
\hline
System & Format & Size & Records & Elapsed time & Elapsed time \\
& & & & no cached data & w/ cached data \\
\hline
DAS (DBS) & XML & 187MB & 368716 & 53 sec & 0.92 sec \\
DAS (PhEDEx) & XML & 164MB & 175884/170131 & 160 sec & 0.94 sec \\
DAS (aggregate) & JSON & 357MB & 374468 & 256 sec & 44.8 sec \\
\hline
\hline
\end{tabular}
\caption{Time required to fetch, parse and aggregate block information
from DBS and PhEDEx systems. In the case of PhEDEx
system we show total number of fetched records 175884 together with
number of records merged in DAS 170131. The total number of DAS records 
were calculated as number of aggregated records plus left overs,
the non-matched records from both systems.}
\label{DAS_benchmark}
\end{table*}

The initial DAS implementation was done in Python language \cite{Python}
and all tests were performed on x64 Linux platform where
1 CPU was used by DAS server and 1 CPU by MongoDB database server. 
%The CPU utilization
%of DAS server was 67\% in average and RAM usage was below 35MB. 
All tests were performed multiple times to avoid fluctuations. 
The elapsed time measured during those tests consists of the 
following componenets:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here the {\it retrieval time} is a time required to access data from remote data-service,
{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
is a time required to convert data-service meta-data to DAS records for every object
we parse, {\it cache insertion} and {\it indexing time} represents time spent on DAS 
cache back-end, {\it aggregation time} is required to merge objects into DAS records based
on their common key (\verb+block.name+ in this case) and {\it output creation time}
is a time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, were only applied to
one of the sub-system, either DBS or PhEDEx, depending on insertion order.
% since the first one was injected into
%the fresh database, while the second required the merging step.

@@

The individual tests of DAS components (DBS and PhEDEx) shown roughly 
the same performance. Although the time of PhEDEx component was bigger 
due to longer retrieval and parsing time.
The time of PhEDEx component shown in Table \ref{DAS_benchmark} is divided 
among retrieval and parsing, $\sim$110 sec, and merging steps, $\sim$50 sec,
respectively. As shown in Table \ref{DAS_benchmark},
the time spent to access individual components (last column) was quite
reasonable, roughly 1 second for both systems. While final time to
get DAS records on disk was about 45 seconds. This was mostly due to I/O and
conversion operations between binary data format used by MongoDB, BSON, to DAS
data format, JSON.

Based on results of stress tests we measured that we can achieve throughput of
7000 docs/sec for insertion, 3400 docs/sec for aggregation and 8500 docs/sec for 
reading operations in DAS, which are suitable for our needs. Since all 
operations were performed on a single CPU a further improvements are expected. 
Also, a necessary tuning can be done by using C++ driver instead of python one.

\section{Summary}
We presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
aggregation layer on top of the existing relational and non-relation data-services
mostly in real time fashion. All of the data were retrieved on demand basis,
while data pre-fetching of most common queries is in our plans. We developed
prototype in python for DAS system which is under commissioning phase right now 
and presented performance studies in this paper. 
The CMS experiment is started to collect its data in December of 2009 and 
we expect to accumulate a few PB of real data each year. 
In addition, the Monte-Carlo samples will be produced at this scale.
Based on test studies performed in CMS, we expect that total size of
meta-data produced each year by experiment will be of the order of
500GB. The DAS should sustain such load and will provide generic 
``data-discovery'' service for CMS experiment in years to come.

@@

\bibitem{Python} Python programming language, http://www.python.org
